---
output: github_document
---


```{r}
knitr::opts_chunk$set(message=FALSE)
```

```{r setup}
library(neonstore)
library(neon4cast) # remotes::install_github("eco4cast/neon4cast", dep=TRUE)
library(tidyverse)
library(tsibble)
library(fable)
```

# A trivial forecast


```{r}
## uuid::UUIDgenerate()
iteration_id <- "c2cb06c2-4bf3-4aa0-9690-1fdd098a7ac5"
```

## Access Target Data

```{r}
targets <-
  "https://data.ecoforecast.org/targets/beetles/beetles-targets.csv.gz" %>% 
  read_csv(col_types = "cDdd") %>% 
  as_tsibble(index = time, key = siteID)
```


```{r}
past <-  targets
forecast_date <- Sys.Date() 

forecast_date <- as.Date("2021-06-01")
past <-  targets %>% filter(time < forecast_date)
```

## Compute a forecast

```{r}
## Compute a simple mean/sd model per site... obviously silly given huge seasonal aspect
null_richness <- past  %>% 
  model(null = MEAN(richness)) %>%
  forecast(h = "1 year")

null_abundance <- past  %>%
  model(null = MEAN(abundance)) %>%
  forecast(h = "1 year")
```


## Visualize the forecast

```{r}
first4 <- unique(null_richness$siteID)[1:4]

null_richness %>% filter(siteID %in% first4)  %>% autoplot(past) + ggtitle("richness")
null_abundance %>% filter(siteID %in% first4)  %>% autoplot(targets) + ggtitle("abundance")
```

## EFI Formatting

EFI requires a flat-file format for forecasts that avoids the use of complex list columns.  
To convey uncertainty, forecasts must be expressed either by giving mean and standard deviation (for predictions that are normally distributed) or must express forecasts as an ensemble of replicate draws from forecast distribution.
The helper function `efi_format()` handles this transformation.

```{r}
## Combine richness and abundance forecasts. drop the 'model' column
f1_forecast <- inner_join(efi_format(null_richness), 
                            efi_format(null_abundance)) %>%
  select(!.model) # we can have only one model per file in EFI format
```

```{r}
forecast_file <- glue::glue("{theme}-{date}-{team}.csv.gz",
                            theme = "beetles", 
                            date=forecast_date,
                            team = "cb_f1")
write_csv(f1_forecast, forecast_file)
```


```{r}
neon4cast::forecast_output_validator(forecast_file)
```


```{r}
yaml_meta <- neon4cast::create_model_metadata(forecast_file)
eml_file <- neon4cast::write_metadata_eml(forecast_file, 
                                          yaml_meta, 
                                          forecast_issue_time = forecast_date, 
                                          forecast_iteration_id = iteration_id)

```


```{r}
submit(forecast_file, metadata = eml_file)
```
